Predicting the Quality of CDR-H3 Antibody Loop Structural Models

### 2022-12-16 paper.tex
Fixed some LaTeX errors

Our ML-model was trained on different feature sets selected using manual
selection as well as algorithmic selection strategies (see methods), in order to
determine the most effective feature selection method. None of the
feature selection methods was a best fit for all ML-models. To create a ML-model implementing the encoding and feature selection strategies best suited for the specific ML-structure, a number of different combinations were tested, summarized in \ref{fig:results_table}. Additional ML-Models were discarded due to poor performance. 
\begin{table}[]
\fontsize{12}{16}\selectfont
  \centering
  \caption{Summary of Machine-Learning Classifier Performances}
  \label{my-label}

@@

\end{table}
\caption{table caption}

@@

    \begin{tabular}{| m{5em} | m{20em}| m{20em} |}
    \hline
Feature Name & Description & method of Calculation \\
\hline
\hline
Sequence & Amino acid sequence of the CDR-H3 loop. & Sequence is given in one-letter amino acid codes.\\
\hline
Length & Number of residues in the CDRH3-loop, which is located at residues H95-H102. & The number of residues are counted.\\
\hline
Sequence Identity & Sequence identity of selected template (SeqA) with input loop sequence (SeqB) is determined after sequence alignment. Calculated by abYmod during modelling. & $Identity(SeqA,SeqB) = 100\%\frac{identical residues}{length(alignment)}$\\
\hline
Sequence similarity & Sequence similarity of selected template (SeqA) with input loop sequence (SeqB) is determined after sequence alignment. Calculated by abYmod during modelling. Similar residues are residues that have undergone conservative substitution. &
$Similarity(SeqA,SeqB) = 100\%\frac{identical residues+similar residues}{length(alignment)}$\\
\hline
Loop Protrusion & Distance of loop residue further away from the loop base & Geometrical calculations \protect\ref{fig:loopdist} \\
\hline
Protruding residue & The Amino acid code of the most protruding loop residue & Using the previously determined point furthest away from the loop base, the residue at this coordinate is determined and given as a one-letter amino acid code.\\
\hline
Charge & Total charge of the loop & Sum of charges of all residues in loop \\
\hline
Charge difference & Difference in total charge compared to template sequence & Difference between the two summed changes \\
\hline
Hydrophobicity & Mean Hydrophobicity values of loop & Based on Eisenberg consensus values\\
\hline
Hydrophobicity  difference & Sum of absolute differences between loop sequence and template loop& Based on Eisenberg consensus values\\
\hline
Accessibility & Total and average accessibility for the loop& lee-Richards method implemented using the pdbsolv method from the BiopTools library.\\
\hline
Side-chain Accessibility & Total and average side-chain accessibility for the loop& lee-Richards method implemented using the pdbsolv method from the BiopTools library.\\
\hline
Relative Accessibility & Total and average relative accessibility for the loop& lee-Richards method implemented using the pdbsolv method from the BiopTools library.\\
\hline
relative side-chain Accessibility & Total and average relative side-chain accessibility for the loop& lee-Richards method implemented using the pdbsolv method from the BiopTools library.\\
\hline
Happiness & Happiness score, taking accessibility and hydropobicity into account. If a residue is 'happy' it will not be a buried hydrophilic or a surface hydrophobic residue. & Hydrophobicity values (see above) are normalized to a range of -1 to +1. Mean accessibility values are calculated as above. If hydrophobicity of loop is $<0$:
$Happiness = 1+Hydrophobicity(1-Accessibility)$
Otherwise:
$Happiness = 1-(Hydrophobicity Accessibility)$\\
\hline
Nr. of Contacts & Nr of contacts made by the residue of the loop within a range of 3.5\AA. Includes mainchain as well as sidechain atoms. Contacts made with residue within and outside of the loop are counted separately and as total. The ratio of inside vs outside is also calculated. & Modified version of the rangecontacts method in the BiopTools library. \\
\hline
Energy & Potential energy of the model. & Calculated by Gromacs during energy minimization step in abYmod modelling. \\
\hline
Lowest BLOSUM 62 Scoring Residue Pair & Each possible residue pair in the CDR-H3loop is scored by their BLOSUM 62 score. The lowest scoring pair's BLOSUM62 value will be combined with their residue separation to form the metric. & With separation being hte nuber of residues between the worst residue pair, and the worst score being the lowest BLOSUM62 score achieved by a residue pair, the metric is calculated as follows: $WorstBLOSUM = -log_2(separation)(worst score)$
\hline

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.5]{protrusion.eps}
  \caption {Diagram
    visualizing the process underlying the protrusion
    calculation. First, the base residues (i.e.\ H95 and H102, shown as
    red spheres) of the CDR-H3 (grey circles) are identified. Then, a
    line is drawn between the two \ca\ atoms of these residues.
    The distance of the \ca-atom of each residue in the
    CDR-H3 loop to this line is calculated (d). The residue which has the
    greatest distance to the line is output
    as one-letter amino acid code and used as feature. The distance d in
    \AA\is used as the `protrusion'
    feature.}
  \label{fig:angle}
\end{figure}

>>>>

Our ML-model was trained on different feature sets selected using manual
selection as well as algorithmic selection strategies (see methods), in order to
determine the most effective feature selection method. None of the
feature selection methods was a best fit for all ML-models. To create a ML-model implementing the encoding and feature selection strategies best suited for the specific ML-structure, a number of different combinations were tested, summarized in \ref{fig:results_table}. Additional ML-Models were discarded due to poor performance.

\begin{table}[]
\fontsize{12}{16}\selectfont
  \centering
  \caption{Summary of Machine-Learning Classifier Performances}
  \label{my-label}

@@

  \end{adjustbox}
\end{table}

@@

    \begin{tabular}{| m{5em} | m{20em}| m{20em} |}
    \hline
Feature Name & Description & method of Calculation \\
\hline
\hline
Sequence & Amino acid sequence of the CDR-H3 loop. & Sequence is given in one-letter amino acid codes.\\
\hline
Length & Number of residues in the CDRH3-loop, which is located at residues H95-H102. & The number of residues are counted.\\
\hline
Sequence Identity & Sequence identity of selected template (SeqA) with input loop sequence (SeqB) is determined after sequence alignment. Calculated by abYmod during modelling. & $Identity(SeqA,SeqB) = 100\%\frac{identical residues}{length(alignment)}$\\
\hline
Sequence similarity & Sequence similarity of selected template (SeqA) with input loop sequence (SeqB) is determined after sequence alignment. Calculated by abYmod during modelling. Similar residues are residues that have undergone conservative substitution. &
$Similarity(SeqA,SeqB) = 100\%\frac{identical residues+similar residues}{length(alignment)}$\\
\hline
Loop Protrusion & Distance of loop residue further away from the loop base & Geometrical calculations \protect\ref{fig:loopdist} \\
\hline
Protruding residue & The Amino acid code of the most protruding loop residue & Using the previously determined point furthest away from the loop base, the residue at this coordinate is determined and given as a one-letter amino acid code.\\
\hline
Charge & Total charge of the loop & Sum of charges of all residues in loop \\
\hline
Charge difference & Difference in total charge compared to template sequence & Difference between the two summed changes \\
\hline
Hydrophobicity & Mean Hydrophobicity values of loop & Based on Eisenberg consensus values\\
\hline
Hydrophobicity  difference & Sum of absolute differences between loop sequence and template loop& Based on Eisenberg consensus values\\
\hline
Accessibility & Total and average accessibility for the loop& lee-Richards method implemented using the pdbsolv method from the BiopTools library.\\
\hline
Side-chain Accessibility & Total and average side-chain accessibility for the loop& lee-Richards method implemented using the pdbsolv method from the BiopTools library.\\
\hline
Relative Accessibility & Total and average relative accessibility for the loop& lee-Richards method implemented using the pdbsolv method from the BiopTools library.\\
\hline
relative side-chain Accessibility & Total and average relative side-chain accessibility for the loop& lee-Richards method implemented using the pdbsolv method from the BiopTools library.\\
\hline
Happiness & Happiness score, taking accessibility and hydropobicity into account. If a residue is 'happy' it will not be a buried hydrophilic or a surface hydrophobic residue. & Hydrophobicity values (see above) are normalized to a range of -1 to +1. Mean accessibility values are calculated as above. If hydrophobicity of loop is $<0$:
$Happiness = 1+Hydrophobicity(1-Accessibility)$
Otherwise:
$Happiness = 1-(Hydrophobicity Accessibility)$\\
\hline
Nr. of Contacts & Nr of contacts made by the residue of the loop within a range of 3.5\AA. Includes mainchain as well as sidechain atoms. Contacts made with residue within and outside of the loop are counted separately and as total. The ratio of inside vs outside is also calculated. & Modified version of the rangecontacts method in the BiopTools library. \\
\hline
Energy & Potential energy of the model. & Calculated by Gromacs during energy minimization step in abYmod modelling. \\
\hline
Lowest BLOSUM 62 Scoring Residue Pair & Each possible residue pair in the CDR-H3loop is scored by their BLOSUM 62 score. The lowest scoring pair's BLOSUM62 value will be combined with their residue separation to form the metric. & With separation being hte nuber of residues between the worst residue pair, and the worst score being the lowest BLOSUM62 score achieved by a residue pair, the metric is calculated as follows: $WorstBLOSUM = -log_2(separation)(worst score)$\\
\hline

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.5]{protrusion.eps}
  \caption {Diagram
    visualizing the process underlying the protrusion
    calculation. First, the base residues (i.e.\ H95 and H102, shown as
    red spheres) of the CDR-H3 (grey circles) are identified. Then, a
    line is drawn between the two \ca\ atoms of these residues.
    The distance of the \ca-atom of each residue in the
    CDR-H3 loop to this line is calculated ($d$). The residue which has the
    greatest distance to the line is output
    as one-letter amino acid code and used as feature. The distance $d$ in
    \AA\ is used as the `protrusion'
    feature.}
  \label{fig:angle}
\end{figure}
